---
  title: Build synthetic networks
  author: Giulia Bertagnolli
  output:
    html_document:
      theme: paper
      fig_caption: true
      fig_align: 'center'
      code_folding: hide
      toc: true
      toc_depth: 3
      toc_float: true
      number_sections: true
    github_document:
      pandoc_args: --webtex
  bibliography: biblio.bib
---

```{r load-packages, echo = FALSE, warning = FALSE}
rm(list = ls())
library(igraph)
library(tidyverse)
library(ggstance)
```

```{r setup, echo = FALSE}
knitr::opts_chunk$set(dev = c("png", "pdf"),
                      fig.path = "imgs/synthetic/",
                      fig.width = 4, fig.height = 3)
igraph_options(vertex.label = NA, vertex.frame.color = "white",
               vertex.frame.size = 3, edge.arrow.size = .5)
```

```{r setup-sim, echo = F}
set.seed(37)  # seed for generation of random samples
n <- 30       # number of samples
```

## Full network topology (clique)


```{r synthetic, echo = TRUE}
N <- 30
g <- make_full_graph(n = N)
m <- gsize(g)
```

We start creating the topology of the networks, a full network with `r N` and `r m` edges. Upon this naive topology  we impose edge weights sampled from two distributions:

* Poisson with varying parameter $\lambda$ for homogeneous weights,
* power-law with varying parameter $\alpha$ for heterogeneous weights.

We henceforth indicate by $W$ the random variable representing the edge weight, i.e. the intensity of an interaction.

### Homogeneous weights distribution

$W \sim Pois(\lambda)$ with $\lambda \in \{1, \dots, 12\}$.
With `rpois(m, lambda)` we take a sample of size `r m` from the Poisson distribution with parameter `lambda`.

```{r setup-pois, echo = TRUE}
lambdas <- 1:12
```

```{r generate-pois, echo = TRUE}
weights_pois <- lapply(lambdas, function(lambda) {
  sapply(1:n, function(i) {
    1 + rpois(m, lambda)
  })
});
```


```{r g-pois, echo = F}
g_pois <- lapply(seq_along(lambdas), function(l) {
  tmp <- g
  E(tmp)$weight <- weights_pois[[l]][, 1]
  return(tmp)
})
```

```{r save-pois, echo = FALSE}
save(weights_pois, file = "data/synthetic/weights_pois.RData")
save(g_pois, file = "data/synthetic/g_pois.RData")
```

```{r weights-pois-boxplots, echo = FALSE, fig.width = 4, fig.height = 3}
# customise plot theme
my_bp_theme <- theme(
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.line.x = element_blank(),
  axis.line.y = element_line(color = "grey")
)
# create and save plots in a loop
bps_pois <- lapply(seq_along(weights_pois), function(i) {
  tibble(
    "weight" = as.numeric(weights_pois[[i]])
  ) %>%
    ggplot(aes(x = weight, y = -.1)) +
      # boxplot of the samples
      geom_boxplot(width = .1, na.rm = T) +
      geom_hline(yintercept = 0, color = "grey") +
      annotate("text", x = seq(5, 20, 5), y = rep(-.02, 4),
        label = seq(5, 20, 5), color = "grey20", size = 3) +
      # pmf of the poisson
      geom_point(data = tibble(
        "x" = 1:26, "y" = dpois(0:25, lambda = lambdas[i])), aes(x, y), inherit.aes = FALSE) +
      scale_x_continuous(limits = c(1, 26), expand = c(.01, .01)) +
      scale_y_continuous(limits = c(-.17, .4), breaks = c(0, .2, .4)) +
      labs(y = "", x = "edge wieght") +
      my_bp_theme -> p
      show(p)
      return(p)
});
```

---

### Heterogeneous weights distribution

$W \sim p(w; \alpha, w_{\min})$ that is:
$$
p(w; x_{\min}, \alpha) = \frac{\alpha-1}{w_{\min }}\left(\frac{w}{w_{\min }}\right)^{-\alpha}
$$
see @clauset2009power for a reference.

```{r setup-pwl, echo = TRUE}
# power-law from uniformly randomly sampled (Clauset Appx.D)
rpower <- function(n, x_min, alpha) {
  r <- runif(m)
  x_min * (1 - r) ** (-1 / (alpha - 1))
}
dpower <- function(x, x_min, alpha) {
  (alpha - 1) / x_min * ((x / x_min) ^ -alpha)
}
# parameter space
x_min <- 5
alphas <- seq(1.5, 7, .5)
```

```{r generate-pwl, echo = TRUE}
weights_pwl <- lapply(alphas, function(alpha) {
  sapply(1:n, function(i) {
    rpower(n, x_min, alpha)
  });
})
```

```{r g-pwl, echo = FALSE}
g_pwl <- lapply(seq_along(alphas), function(l) {
  tmp <- g
  E(tmp)$weight <- weights_pwl[[l]][, 1]
  return(tmp)
})
```

```{r save-pwl, echo = FALSE}
save(g_pwl, file = "data/synthetic/g_pwl.RData")
save(weights_pwl, file = "data/synthetic/weights_pwl.RData")
```

```{r weights-pwl-boxplots, echo = FALSE, fig.width = 4, fig.height = 3}
bps_pwl <- lapply(seq_along(weights_pwl), function(i) {
  pwl_pdf <- tibble(
    "x" = seq(5, 20, .1),
    "px" = dpower(seq(5, 20, .1), x_min = 5,  alpha = alphas[i])
  )
  tibble(
    "weight" = as.numeric(weights_pwl[[i]])
  ) %>%
    ggplot(aes(x = weight, y = -max(pwl_pdf$px) / 8)) +
      geom_boxploth(width = (max(pwl_pdf$px) / 10), na.rm = T) +
      geom_hline(yintercept = 0, color = "grey") +
      annotate("text", x = seq(5, 20, 5), y = rep(-max(pwl_pdf$px) / 100, 4),
        label = seq(5, 20, 5), color = "grey20", size = 3) +
      geom_line(data = pwl_pdf, aes(x, px), inherit.aes = FALSE, na.rm = T) +
      scale_x_continuous(limits = c(5, 20), expand = c(.01, .01)) +
      scale_y_continuous(breaks = seq(0, max(pwl_pdf$px), length.out = 5)) +
      labs(y = "", x = "edge wieght") +
      my_bp_theme -> p
      show(p)
      return(p)
});
```

## Weighted networks from network models

The following code does **not** generate the same data as in our paper @bertagnolli2020quantifying, which are available under reasonable request.
Here we provide some sample code, for obtaining **one** realisation of each network model topology and then generate their edge weights as in @bertagnolli2020quantifying.

The parameters used to generate the networks topology have been chosen such that for $N=256$ we get $\langle k \rangle=12$.

### Barabasi--Albert model with weights

```{r synthetic-BA}
N <- 256
g_BA <- sample_pa(n = N, m = 12 / 2, power = 1, directed = FALSE)
```

This network is unweighted and we add link weights following the rule
$$w_{ij} = k_i^{- \beta}$$
for varying parameter $\beta$.

```{r setup-parameters-weights}
betas <- seq(-5, 5)
unitary_matrix <- matrix(1, N, N)
```

#### Generating links weights

```{r BA-degrees}
degree_BA <- degree(g_BA)
```

```{r BA-weights}
weighted_adj_BA_list <- lapply(betas, function(beta) {
  degree_BA ^ beta * unitary_matrix * as_adjacency_matrix(g_BA)
})
```

Now, we have a list (where the index runs from 1 to `r length(betas)` and represents the `beta` parameter), where each element of the list is a weighted adjacency matrix.
Observe that the 11th element corresponds to $\beta = 0$ and to the unweighted case. Except for the case $\beta = 0$, the networks become **directed**, since the weighted adjacency matrices are no more symmetric.


```{r weighted-BAs}
weighted_g_BA_list <- lapply(seq_along(betas), function(i) {
  W <- weighted_adj_BA_list[[i]]
  graph_from_adjacency_matrix(W, weighted = TRUE)
});
```

```{r plot-BA, echo = FALSE, include = TRUE}
wBA_plots <- lapply(weighted_g_BA_list, function(g) {
  edge_width <- E(g)$weight / max(E(g)$weight)
  plot(g, edge.width = 1 + edge_width);
})
```

### Watts--Strogatz model with weights

We do the same for a WS model.

```{r synthetic-WS, echo = TRUE}
g_WS <- sample_smallworld(dim = 2, size = sqrt(N), nei = 2, p = 0.1,
                          multiple = F, loops = F)
```

#### Generating links weights

```{r WS-degrees}
degree_WS <- degree(g_WS)
```

```{r WS-weights}
weighted_adj_WS_list <- lapply(betas, function(beta) {
  degree_WS ^ beta * unitary_matrix * as_adjacency_matrix(g_WS)
})
```

```{r weighted-WSs}
weighted_g_WS_list <- lapply(seq_along(betas), function(i) {
  W <- weighted_adj_WS_list[[i]]
  g <- graph_from_adjacency_matrix(W, weighted = TRUE)
  return(g)
});
```

```{r plot-WS, echo = FALSE, include = TRUE}
wWS_plots <- lapply(weighted_g_WS_list, function(g) {
  edge_width <- E(g)$weight / max(E(g)$weight)
  plot(g, edge.width = 1 + edge_width);
});
```

```{r save-BA-WS, echo = FALSE}
save(g_BA, degree_BA, weighted_adj_BA_list, weighted_g_BA_list,
     file = "data/synthetic/weighted_WS.RData")
save(g_WS, degree_WS, weighted_adj_WS_list, weighted_g_WS_list,
     file = "data/synthetic/weighted_WS.RData")
```
